\documentclass[11pt]{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{xcolor}
\usepackage{float}
\usepackage{listings}
\usepackage{xcolor}
\usepackage[mathscr]{eucal}
\usepackage[most]{tcolorbox}

\DeclareTColorBox{notebox}{ O{Note} O{blue} }{
  colback=#2!5!white,
  colframe=#2!75!black,
  fonttitle=\bfseries,
  title={#1}
}


\lstset{
    % change caption to be below the listing
    % choose appropriate font style.
    captionpos=b,
  basicstyle=\footnotesize\ttfamily,
  numbers=left,
  numberstyle=\tiny,
  stepnumber=1,
  numbersep=8pt,
  frame=single,
  breaklines=true,
  tabsize=2,
  showstringspaces=false,
  keywordstyle=\color{blue},
  commentstyle=\color{gray},
  stringstyle=\color{teal},
}

\title{Radiation-Hydrodynamics Simulation}
\author{Tal Ramot}
\date{Exams week 3, 2026}

\begin{document}

\maketitle

\section{Hydrodynamics scheme}
The scheme uses the principle of operator splitting, where the hydrodynamics and radiation are solved separately. The hydrodynamics is solved using a finite volume Lagrangian method (e.g. Richtmyer-Morton) and the radiation is solved using a finite difference method. The coupling between the two is done through the source terms in the hydrodynamics equations, which are updated after each radiation step.
\subsection{Hydrodynamics step}
The following uses the Euler equations:
\begin{gather}
  \frac{\partial u}{\partial t} = - \frac{\partial P}{\partial m} \\
  \frac{\partial V}{\partial t} = \frac{\partial u}{\partial m} \\
  \frac{\partial e}{\partial t} = -P \frac{\partial u}{\partial m} \\
  p = r \rho e \\
  P = p + q
\end{gather}
where $E_{\text{rad}}$ is the energy exchange with the radiation field, which is updated after the radiation step (kept aside for now). 
In the lagrangian scheme, we begin with a certain pressure that applies a force on the mass elements, which causes them to accelerate. The scheme goes as follows:
\begin{enumerate}
  
  \item Compute acceleration from the pressure gradient: 
  \begin{equation}
    a_i^{n} = -2 \cdot \frac{\left(p_{i+\frac{1}{2}}^n + q_{i+\frac{1}{2}}^n\right) - \left(p_{i-\frac{1}{2}}^n + q_{i-\frac{1}{2}}^n\right)}{m_{i+\frac{1}{2}} + m_{i-\frac{1}{2}}}
  \end{equation}
  \item Update velocity using the acceleration:
  \begin{equation}
    u_i^{n+\frac{1}{2}} = u_i^{n} + a_i^{n} \Delta t
  \end{equation}
  \item Update node positions using the updated velocities (which is exactly a helper step for updating the volume):
  \begin{equation}
    x_i^{n + 1} = x_i^{n} + u_i^{n+\frac{1}{2}} \Delta t
  \end{equation}
  \item Update the volume and density of each cell using the new node positions:
  \begin{align}
    V_{i+\frac{1}{2}}^{n+1} &= \frac{x_{i+1}^{n+1} - x_i^{n+1}}{m_{i+\frac{1}{2}}} \\
    \rho_{i+\frac{1}{2}}^{n+1} &= \frac{1}{V_{i+\frac{1}{2}}^{n+1}}
  \end{align}
  \item Calculate artificial viscosity $q$ to handle shocks, using the von Neumann-Richtmyer prescription:
  \begin{equation}
    q_{i+\frac{1}{2}}^{n+1} = \begin{cases}
      \sigma \rho_{i+\frac{1}{2}}^{n+1} \left(u_{i+1}^{n+1} - u_i^{n+1}\right)^2 & \text{if } u_i^{n+1} > u_{i+1}^{n+1} \\
      0 & \text{otherwise}
    \end{cases}
  \end{equation}
  \item Calculate the new spcific internal energy using the updated pressure and velocity:
  \begin{equation}
    e_{i+\frac{1}{2}}^{n+1} = \frac{e_{i+\frac{1}{2}}^{n} -\frac{1}{2} \left(p_{i+\frac{1}{2}}^{n} + q_{i+\frac{1}{2}}^{n} + q_{i+\frac{1}{2}}^{n+1}\right) \left(\frac{V_{i+\frac{1}{2}}^{n+1} - V_{i+\frac{1}{2}}^{n}}{m_{i+\frac{1}{2}}}\right)}{1 + \frac{\gamma - 1}{2} \rho_{i+\frac{1}{2}}^{n+1} \frac{V_{i+\frac{1}{2}}^{n+1} - V_{i+\frac{1}{2}}^{n}}{m_{i+\frac{1}{2}}}}
  \end{equation}
  \item Calculate the new pressure using the equation of state:
  \begin{equation}
    p_{i+\frac{1}{2}}^{n+1} = (\gamma - 1) \rho_{i+\frac{1}{2}}^{n+1} e_{i+\frac{1}{2}}^{n+1}
  \end{equation}
  \item Compute acceleration from the \textbf{new} pressure gradient: 
  \begin{equation}
    a_i^{n+1} = -2 \cdot \frac{\left(p_{i+\frac{1}{2}}^{n+1} + q_{i+\frac{1}{2}}^{n+1}\right) - \left(p_{i-\frac{1}{2}}^{n+1} + q_{i-\frac{1}{2}}^{n+1}\right)}{m_{i+\frac{1}{2}} + m_{i-\frac{1}{2}}}
  \end{equation}
  \item Update velocity using the new acceleration:
  \begin{equation}
    u_i^{n+1} = u_i^{n+\frac{1}{2}} + a_i^{n+1} \frac{\Delta t}{2}
  \end{equation}
  and go back to step 2 for the next iteration. \textcolor{red}{Make sure to understand the reason for the half time step in the velocity update - it makes the scheme second-order and less noisy}.
\end{enumerate}
\subsection{Coupling with radiation}
This is all good for the hydrodynamics step, but we need to couple it with radiation! After step 9, we define:
\begin{equation}
  e^{*} = e^{n+1} 
\end{equation}
which means that the energy isn't updated yet! we then assume that $U_m^{n} = \rho e^{*}$ is the energy density of the matter.

\section{Radiation Scheme}
\subsection{Physical PDEs}
\begin{enumerate}
  \item Writing the diffusion equation for the radiation energy $E=aT_r^4$ and the effective material radiative energy $U_R = aT_m^4$ (the energy that the material, modeled as a black body, would radiate outwards):
  \begin{equation}
      \frac{\partial{E}}{\partial t}+\rho \frac{\partial F}{\partial m} = \chi c\sigma(U_R-E) 
  \end{equation}
  \begin{equation}
      \frac{\partial U_m}{\partial t}=-\chi c\sigma (U_R-E)\Rightarrow
      \frac{1}{\beta} \frac{\partial U_R}{\partial t}=-\chi c\sigma (U_R-E)
  \end{equation}
  \begin{equation}
      F=-D\rho\frac{\partial E}{\partial m}\;\;\;\;\left(D=\frac{c}{3\sigma(T,\rho)}\right)
  \end{equation}
  In the equation here, one assumes that during the radiation substep we keep $\rho^*$ (to be defined later) fixed, and then linearize $U_R(U_m)$ locally around $U_m^*$ to get the relation $\frac{1}{\beta} \frac{\partial U_R}{\partial t}=-\chi c\sigma (U_R-E)$, where $\beta = \frac{dU_R}{dU_m}|_{U_m^*}$ is the coupling coefficient between the radiation and matter that satifies $dU_R \approx \beta dU_m$ around $U_m^*$.
  \item The self similar relations for the opacity (and thus the diffusion constant):
  \begin{equation}
    \frac{1}{\kappa_R} = g T^\alpha \rho^{-\lambda}
  \end{equation}
  \begin{equation}
    \frac{1}{\sigma(T,\rho)} = \frac{1}{\rho\kappa_R(T)} = g T^\alpha \rho^{-\lambda - 1} \Rightarrow D = \frac{c}{3} g T^\alpha \rho^{-\lambda - 1}
  \end{equation}
  and the self similar relation for the internal energy:
  \begin{equation}
      e = f T^\gamma \rho^{-\mu}
  \end{equation}
  % alarm bells icon at the beginning of the line
  where $e$ is the specific internal energy, which means the material energy density is:
  \begin{equation}
      U_m(T) = \rho e = f T^\gamma \rho^{-\mu + 1}
  \end{equation}
  \item Notice that we've converted the space derivatives to mass derivatives, using the relation:
  \begin{equation}
    \frac{\partial}{\partial x} = \frac{\partial m}{\partial x} \frac{\partial}{\partial m} = \rho \frac{\partial}{\partial m}
  \end{equation}
  as done in the Appendix.
  \item To sum up, we have \textbf{a system of PDEs over $T_m$ and $T_R$}:
  \begin{equation}
      \frac{\partial{\left(aT_r^4\right)}}{\partial t}-\rho \frac{\partial}{\partial m}\left(D \rho \frac{\partial (aT_r^4)}{\partial m}\right) = \chi c\sigma a(T_m^4-T_r^4) 
  \end{equation}
  \begin{equation}
      \frac{1}{\beta} \frac{\partial \left(aT_m^4\right)}{\partial t}=-\chi c\sigma a(T_m^4-T_r^4)
  \end{equation}
  The system is solved over a domain of $N$ cells with BC of ... \textcolor{red}{to be defined}.
\end{enumerate}
\subsection{Finite Difference Scheme}
\subsubsection{Discretization of a lagrangian PDE}
\begin{enumerate}
  \item In the diffusion equation, we have terms like:
  \begin{equation}
    \frac{\partial E}{\partial x} = \rho \frac{\partial E}{\partial m}
  \end{equation}
  \textbf{but how are they being discretized?}
  We can discretize them by simply writing (\textcolor{blue}{leaving $\rho$ outside of discussion for now}):
  \begin{equation}
    \frac{\partial E}{\partial m} \approx \frac{E_{i+\frac{1}{2}}^{n} - E_{i-\frac{1}{2}}^{n}}{\Delta m_i}
  \end{equation}
  \begin{equation}
    \Delta m_i := \frac{1}{2} \left((m_{i+1} - m_{i}) + (m_{i} - m_{i-1})\right) = \frac{1}{2} (\Delta m_{i+\frac{1}{2}} + \Delta m_{i-\frac{1}{2}})
  \end{equation}
  where the lagrangina coordinate $m_i$ for $i=0,...,N$ is defined as the mass of the material up to the vertex located at $x_i$. Therefore $m_i$ = $\sum_{j = 0}^{i} \Delta m_{i+\frac{1}{2}}$ where $\Delta m_{i+\frac{1}{2}} = m_{i+1} - m_{i}$ is the mass of the cell $j = i+\frac{1}{2}$ ($j = 1,...,N$) between the vertices $i$ and $i+1$. 
  \item It may also be denoted $\Delta m_i = m_{i+\frac{1}{2}} - m_{i-\frac{1}{2}}$ where $m_{i+\frac{1}{2}} = m_{i} + \frac{1}{2} \Delta m_{i+\frac{1}{2}}$ and $m_{i-\frac{1}{2}} = m_{i-1} + \frac{1}{2} \Delta m_{i-\frac{1}{2}}$.
    % note: \Delta m_i is the mass of the cell i, and \Delta m_{i+1/2} is the mass difference between vertices i and i+1.

  \begin{equation}
    \begin{aligned}
    \Delta m_i &= m_{i+\frac{1}{2}} - m_{i-\frac{1}{2}} \\ 
    &= \left(m_{i} + \frac{1}{2} \Delta m_{i+\frac{1}{2}}\right) - \left(m_{i-1} + \frac{1}{2} \Delta m_{i-\frac{1}{2}}\right) = \frac{1}{2} \left(\Delta m_{i+\frac{1}{2}} + \Delta m_{i-\frac{1}{2}}\right)
    \end{aligned}
  \end{equation}
  \item One should also regard the problem when taking the second derivative of $E$ with respect to $m$ - rising from the first derivative of the flux!
  \begin{equation}
    \frac{\partial F}{\partial x} = \rho \frac{\partial F}{\partial m} = \rho \frac{\partial}{\partial m} \left( -D\rho \frac{\partial E}{\partial m} \right)
  \end{equation}
  \textbf{What a mess! How are we supposed to discretize this?}
  The solution is to discretize step by step - first the discretize the first derivative of the flux, and then the first derivative of each $E$ term.
  \begin{equation}
    \frac{\partial F}{\partial m} \approx \frac{F_{i+1}^{n} - F_{i}^{n}}{\Delta m_{i + \frac{1}{2}}}
  \end{equation}
  and it makes sense because the flux is defined at the faces of the cells $i+\frac{1}{2}$ whose mass is actually $\Delta m_{i+\frac{1}{2}} = m_{i+1} - m_{i}$.
  Then, one should discretize each of the flux terms, as we have already done for the first derivative of $E$:
  \begin{equation}
    \begin{gathered}
      F_{i+1}^{n} = -\rho_{i+1}^{n} D_{i+1}^{n} \frac{E_{i+\frac{3}{2}}^{n} - E_{i+\frac{1}{2}}^{n}}{\Delta m_{i+1}} \\
      F_{i}^{n} = -\rho_{i}^{n} D_{i}^{n} \frac{E_{i+\frac{1}{2}}^{n} - E_{i-\frac{1}{2}}^{n}}{\Delta m_{i}}
    \end{gathered}
  \end{equation}
  where $\Delta m_i$, $\rho_i$ and $D_i$ terms are all explained in the previous bullets.
  \item \textcolor{blue}{NOTE that in our case $\Delta m_{i+\frac{1}{2}}$ is constant for all $i$ and therefore $\Delta m_i = const$ and is a \textbf{predefined parameter} of the simulation.}
  \subsubsection{Discretization of $D$ and $\rho$ faces}
  Since the derivative is a property of the vertex located between two adjacent cells, we should also pay attention to the fact that $F$, $D$ and $\rho$ should be defined at the faces of the cells (e.g. the vertices).

  Since the natural definition of $D$ and $\rho$ is at the cell centers (we are used to terms like $\rho_{i+\frac{1}{2}}^{n}$ and $D_{i+\frac{1}{2}}^{n}$), we need to define them at the faces of the cells using some kind of averaging.

  \item The most common way to average $D$ and $\rho$ at the faces is to use the average mean:

  \begin{equation}
    \begin{aligned}
    D_{i+\frac{1}{2}} = \frac{D_{i} + D_{i+1}}{2} \\
    \rho_{i+\frac{1}{2}} = \frac{\rho_{i} + \rho_{i+1}}{2}
    \end{aligned}
  \end{equation}
  \item In our simulation, the constant \texttt{HARMONIC\_MEAN} is set by default to \texttt{False}, which means that we use the arithmetic mean to average $D$ and $\rho$ at the faces.



  \subsubsection{Solution of the discretized system of equations}
  \item We've seen that one can \textbf{eliminate the effective radiation density $U_R$} from the finite-difference equation for the matter energy. The discretized equation for the material energy density is:
  \begin{equation}
      \frac{1}{ \beta_{i+\frac{1}{2}}^{n}}\frac{[U_R]_{i+\frac{1}{2}}^{n+1} - [U_R]_{i+\frac{1}{2}}^n}{\Delta t} = \chi c \sigma_{i+\frac{1}{2}}^{n} (E_{i+\frac{1}{2}}^{n+1} - [U_R]_{i+\frac{1}{2}}^{n+1})
  \end{equation}
  \begin{equation}
      (1 +  \beta_{i+\frac{1}{2}}^{n} \Delta t \chi c \sigma_{i+\frac{1}{2}}^{n}) [U_R]_{i+\frac{1}{2}}^{n+1} =  \beta_{i+\frac{1}{2}}^{n}\Delta t \chi c \sigma_{i+\frac{1}{2}}^{n} E_{i+\frac{1}{2}}^{n+1} + [U_R]_{i+\frac{1}{2}}^n
  \end{equation}
  We want to solve for $[U_R]_{i+\frac{1}{2}}^{n+1}$. First, we can denote $A_{i+\frac{1}{2}}^n = \beta_{i+\frac{1}{2}}^{n} \Delta t \chi c \sigma_{i+\frac{1}{2}}^{n}$ as the coupling coefficient between the radiation and matter, which gives us:
  \begin{equation}
      (1 + A_{i+\frac{1}{2}}^n) [U_R]_{i+\frac{1}{2}}^{n+1} =  A_{i+\frac{1}{2}}^n E_{i+\frac{1}{2}}^{n+1} + [U_R]_{i+\frac{1}{2}}^n
  \end{equation}
  \begin{equation}
      \boxed{[U_R]_{i+\frac{1}{2}}^{n+1} = \frac{A_{i+\frac{1}{2}}^n}{1 + A_{i+\frac{1}{2}}^n} E_{i+\frac{1}{2}}^{n+1} + \frac{1}{1 + A_{i+\frac{1}{2}}^n}[U_R]_{i+\frac{1}{2}}^n}
  \end{equation}
  \item Furthermore, after subtituting the expressions for $[U_R]_{i+\frac{1}{2}}^{n+1}$ \& $F_i^{n}$ into the discretized equation for the radiation energy $E$, we get a tridiagonal implicit scheme for $E_i^{n+1}$:
  \begin{equation}
    \boxed{a_{i+\frac{1}{2}}^n E_{i-\frac{1}{2}}^{n+1}+b_{i+\frac{1}{2}}^n E_{i+\frac{1}{2}}^{n} +c_{i+\frac{1}{2}}^n E_{i+\frac{3}{2}}^{n+1}=d_{i+\frac{1}{2}}^n}
  \end{equation}
  where the coefficients $a_i^n,b_i^n,c_i^n,d_i^n$ are calculated in the appendix.
  \item We notice that the coefficients $a_{i}^n,b_{i}^n,c_{i}^n,d_{i}^n$ depend on the hydrodynamics variables (e.g. density, temperature) through the diffusion constant $D$ and the opacity $\sigma$. The dependency is given through Rosen's formulae for the opacity and the internal energy. First, one should find the temperature at the previous time step:
  \begin{equation}
      T^n_{i+\frac{1}{2}}=\sqrt[4]{\frac{[U_R]^n_{i+\frac{1}{2}}}{a}}
  \end{equation}
  and then substitute it into the relations for $D(T),\beta(T)$:
  \begin{gather}
      D_{i+\frac{1}{2}}^n = \frac{c}{3} g \left[T_{i+\frac{1}{2}}^n\right]^\alpha \left[\rho_{i+\frac{1}{2}}^n\right]^{-\lambda - 1} \\
      \beta_{i+\frac{1}{2}}^n = \frac{dU_R}{dU_m} = \frac{4a}{f\gamma} \left[T_{i+\frac{1}{2}}^n\right]^{4 - \gamma} \left[\rho_{i+\frac{1}{2}}^n\right]^{\mu - 1} \\
      \frac{1}{\sigma_{i+\frac{1}{2}}^n} = \frac{1}{\rho \kappa_R (T)} = g \left[T_{i+\frac{1}{2}}^n\right]^\alpha \left[\rho_{i+\frac{1}{2}}^n\right]^{-\lambda - 1}
  \end{gather}
  NOTE: In the coupled scheme, we will use the temperature $T_{i+\frac{1}{2}}^*$ computed from the hydrodynamics step \textbf{directly} to calculate $D,\beta,\sigma$ and substitute in the tridiagonal system, no need to go through $U_R$ in the coupled scheme.
  \item To sum up, the complete algorithm for the radiation scheme is as follows:
  \begin{enumerate}
    \item Compute the temperature at the previous time step using the effective radiation density $U_R$.
    \item Compute the diffusion constant $D$, the coupling coefficient $\beta$ and the opacity $\sigma$ using the temperature and density at the previous time step.
    \item Compute the coefficients $a_i^n,b_i^n,c_i^n,d_i^n$ for the tridiagonal system.
    \item Solve the tridiagonal system for $E_i^{n+1}$ using the Thomas algorithm.
    \item Update the effective radiation density $U_R^{n+1}$ using the expression derived above.
  \end{enumerate}
\end{enumerate}
\section{Coupling the two schemes}
\begin{enumerate}
  \item We first perform the hydrodynamics step, which gives us the updated density $\rho^{n+1}$ and specific internal energy $e^{n+1}$. We call it $e^*$ because we haven't updated the energy with the radiation source term yet.
  \item In the radiation scheme, the self-similar profiles for the opacity and internal energy depend on the material temperature. Since we know that $U_m = \rho e^*$, we can compute the temperature at the previous time step using the relation:
  \begin{equation}
    e_{i+\frac{1}{2}}^* = f \left[T_{i+\frac{1}{2}}^*\right]^{\gamma} \left[\rho_{i+\frac{1}{2}}^{*}\right]^{-\mu} \Rightarrow T_{i+\frac{1}{2}}^* = \left(\frac{e_{i+\frac{1}{2}}^*}{f} \left[\rho_{i+\frac{1}{2}}^{*}\right]^{\mu}\right)^{\frac{1}{\gamma}}
  \end{equation}
  \item Apply the radiation scheme using the temperature and density at the previous time step to compute $\beta$, $D$ and $\sigma$ which give us the coefficients for the tridiagonal system, which we solve to get $E_{i+\frac{1}{2}}^{n+1}$ and $\left[U_R\right]_{i+\frac{1}{2}}^{n+1}$. \\
  NOTE: The chat proposes to perform a Picard loop, where we update the temperature using the radiation energy until convergence.
  \item We then update the specific internal energy after the second source term of radiation is applied:
  \begin{equation}
    e_{i+\frac{1}{2}}^{n+1} = \frac{[U_m]_{i+\frac{1}{2}}^{n+1}}{\rho_{i+\frac{1}{2}}^{n+1}}
  \end{equation}
  \item Finally, we update the hydrodynamic pressure using the equation of state:
  \begin{equation}
    p_{i+\frac{1}{2}}^{n+1} = r \rho_{i+\frac{1}{2}}^{n+1} e_{i+\frac{1}{2}}^{n+1}
  \end{equation}
  and go back to the hydrodynamics step for the next iteration!
\end{enumerate}
\section{Corrections}
\subsection{Correcting the scheme with black physics}
\begin{enumerate}
  \item In the radiation diffusion equation, the finite-difference form that we reached previously is:
  \begin{equation}
      LHS = \frac{\partial{E}}{\partial t}+\rho \frac{\partial F}{\partial m} 
    \end{equation}  
  \begin{equation} \Updownarrow \end{equation}
  \begin{equation}
    \begin{gathered}
  LHS = \frac{E_{i+\frac{1}{2}}^{n+1} - E_{i+\frac{1}{2}}^n}{\Delta t} - \frac{\rho_{i+\frac{1}{2}}^{*}}{m_{i+\frac{1}{2}}} \overbrace{\left(\rho_{i+\frac{3}{2}}^n \frac{2D_{i+\frac{1}{2}}^{n} D_{i+\frac{3}{2}}^{n}}{D_{i+\frac{1}{2}}^{n} + D_{i+\frac{3}{2}}^{n}}\frac{E_{i+\frac{3}{2}}^{n+1} - E_{i+\frac{1}{2}}^{n+1}}{m_{i+\frac{3}{2}}}\right)}^{-F_{i+1}^{n+1}} \\ + \frac{\rho_{i+\frac{1}{2}}^{*}}{m_{i+\frac{1}{2}}}\overbrace{\left(\rho_{i-\frac{1}{2}}^n \frac{2D_{i-\frac{1}{2}}^{n} D_{i+\frac{1}{2}}^{n}}{D_{i-\frac{1}{2}}^{n} + D_{i+\frac{1}{2}}^{n}}\frac{E_{i+\frac{1}{2}}^{n+1} - E_{i-\frac{1}{2}}^{n+1}}{m_{i+\frac{1}{2}}}\right)}^{-F_{i}^{n+1}}
  \end{gathered}
  \end{equation}
  but wait! If we assume black physics, the radiation energy should be changed to $E^*$ and not $E^{n+1}$. That's why the new LHS is:
  \begin{equation}
    \begin{gathered}
  LHS = \frac{\textcolor{blue}{E_{i+\frac{1}{2}}^{*}} - E_{i+\frac{1}{2}}^n}{\Delta t} - \frac{\rho_{i+\frac{1}{2}}^{*}}{m_{i+\frac{1}{2}}} \overbrace{\left(\rho_{i+\frac{3}{2}}^n \frac{2D_{i+\frac{1}{2}}^{n} D_{i+\frac{3}{2}}^{n}}{D_{i+\frac{1}{2}}^{n} + D_{i+\frac{3}{2}}^{n}}\frac{E_{i+\frac{3}{2}}^{n+1} - E_{i+\frac{1}{2}}^{n+1}}{m_{i+\frac{3}{2}}}\right)}^{-F_{i+1}^{n+1}} \\ + \frac{\rho_{i+\frac{1}{2}}^{*}}{m_{i+\frac{1}{2}}}\overbrace{\left(\rho_{i-\frac{1}{2}}^n \frac{2D_{i-\frac{1}{2}}^{n} D_{i+\frac{1}{2}}^{n}}{D_{i-\frac{1}{2}}^{n} + D_{i+\frac{1}{2}}^{n}}\frac{E_{i+\frac{1}{2}}^{n+1} - E_{i-\frac{1}{2}}^{n+1}}{m_{i+\frac{1}{2}}}\right)}^{-F_{i}^{n+1}}
  \end{gathered}
  \end{equation}
  
  \item And the RHS remains implicit, so it's the same as before:
  \begin{equation}
    RHS = \chi c\sigma(U_R-E)
  \end{equation}
  \begin{equation} \Updownarrow \end{equation}
  \begin{equation}
    \begin{gathered}
  RHS = \chi c \sigma_{i+\frac{1}{2}}^{n} \left(\frac{A_{i+\frac{1}{2}}^n}{1 + A_{i+\frac{1}{2}}^n} E_{i+\frac{1}{2}}^{n+1} + \frac{1}{1 + A_{i+\frac{1}{2}}^n}[U_R]_{i+\frac{1}{2}}^n - E_{i+\frac{1}{2}}^{n+1}\right) \\
  = \chi c \sigma_{i+\frac{1}{2}}^{n}
  \left(\frac{1}{1 + A_{i+\frac{1}{2}}^n}[U_R]_{i+\frac{1}{2}}^n -\frac{1}{1 + A_{i+\frac{1}{2}}^n} E_{i+\frac{1}{2}}^{n+1}\right) \\
  = \frac{\chi c \sigma_{i+\frac{1}{2}}^{n}}{1 + A_{i+\frac{1}{2}}^n}[U_R]_{i+\frac{1}{2}}^n - \frac{\chi c \sigma_{i+\frac{1}{2}}^{n}}{1 + A_{i+\frac{1}{2}}^n} E_{i+\frac{1}{2}}^{n+1} 
  \end{gathered}
  \end{equation}
  \item Conclusion: The new LHS and RHS define a new tridiagonal system of equations for the radiation energy density $E^*$, which creates a new sparse matrix that needs to be solved.
  \textcolor{blue}{Ask Shay about this conclusion - if it's correct, it requires a new algebric derivation of the tridiagonal constants, and implementation in python.}

  \item CORRECTION: My conclusion is incorrect. $E^n$ should be replaced by \textcolor{blue}{$E^*$} and \textcolor{red}{NOT $E^{n+1}$}. It makes sense because we suppose that the radiation energy should also be changed.
  
  \item \underline{Implementation}: The only coefficient that had previously (prior to the correction) included $E^n$ is $d^n$. The new formula for $d^n$ is:
  \begin{equation}
    \begin{gathered}
      d_{i}^n = \frac{\chi c \sigma_{i+\frac{1}{2}}^{n}}{1 + A_{i+\frac{1}{2}}^n}[U_R]_{i+\frac{1}{2}}^n + \frac{1}{\Delta t}\textcolor{blue}{E_{i+\frac{1}{2}}^*}
    \end{gathered}
  \end{equation}
  or in language of arrays:
  \begin{equation}
    \begin{gathered}
      d_{j}^n = \frac{\chi c \sigma_{j}^{n}}{1 + A_{j}^n}[U_R]_{j}^n + \frac{1}{\Delta t}\textcolor{blue}{E_{j}^*}
    \end{gathered}
  \end{equation}
\end{enumerate}
\subsection{Black (LTE) radiation scheme}
\begin{enumerate}
  \item One can also solve the rad-hydro scheme under the LTE assumption that leads to the following simpler equation:
  \begin{equation}
    \begin{aligned}
        \frac{\partial E}{\partial t} + \frac{\partial F}{\partial x} = Q \\
        C_V \frac{\partial T}{\partial t} = -Q
    \end{aligned}
  \end{equation}
  where we denoted:
  \begin{equation}
      \begin{gathered}
          Q = \Sigma_a^\prime (aT_m^4 - E) \\
          F = -D \frac{\partial E}{\partial x}
      \end{gathered}
  \end{equation}
  \begin{equation} \Downarrow \end{equation}
  \begin{equation}
    \begin{gathered}
    \frac{\partial e}{\partial t} + \frac{\partial F}{\partial x} = 0 \\
    F = -D \frac{\partial e}{\partial x}
    \end{gathered}
  \end{equation}
  The LTE assumption means that $T_m = T_R = T$ and the relation between the energy densities and the temperatures is given by:
  \begin{equation}
      \begin{aligned}
          e = f T^\beta \rho^{-\mu} \quad \text{Rosen's notation} \\
          E = a T^4 \quad \text{Stefan-Boltzmann law}
      \end{aligned}
  \end{equation}
  \item The finite-difference schem for this equation is:
  The starting point is the coupled energy equation in LTE conditions from above:
  \begin{equation}
    \begin{gathered}
    \frac{\partial e}{\partial t} = \frac{\partial}{\partial x}\left(D \frac{\partial e}{\partial x}\right) \\ 
    \frac{\partial e}{\partial t} = \rho \frac{\partial}{\partial m}\left(D \rho \frac{\partial e}{\partial m}\right)
    \end{gathered}
  \end{equation}
  In the Lagrangian framework used here, the cell centers are indexed by $i+\frac{1}{2}$, while the cell faces (vertices) are indexed by $i$. In Lagrangian coordinates, the spatial derivative is written in terms of mass coordinates, using $1/\Delta x = \rho^* / \Delta m$, where $\Delta m$ is the cell mass and $\rho^*$ is the (hydro-updated) density at the cell center.

  The finite-difference (implicit) scheme for the diffusion equation in Lagrangian coordinates reads:
  \begin{equation}
    \begin{gathered}
    LHS = \frac{e_{i+\frac{1}{2}}^{n+1} - e_{i+\frac{1}{2}}^n}{\Delta t} \\
    RHS = \frac{1}{\Delta m_{i+\frac{1}{2}}}
      \left[
        D_{i+1} \rho_{i+1}^* \frac{e_{i+\frac{3}{2}}^{n+1} - e_{i+\frac{1}{2}}^{n+1}}{\Delta m_{i+1}}
        - D_{i} \rho_{i}^* \frac{e_{i+\frac{1}{2}}^{n+1} - e_{i-\frac{1}{2}}^{n+1}}{\Delta m_{i}}
      \right]
    \end{gathered}
  \end{equation}
  where $D_i$ is the diffusion coefficient at face (vertex) $i$, $\Delta m_{i}$ is the mass of the cell between vertices $i-1$ and $i$, and $\rho_{i+\frac{1}{2}}^*$ is the density at the cell center after the hydro update.

  The harmonic mean is still used for $D_i$:
  \begin{equation}
    D_{i} = \frac{2 D_{i-\frac{1}{2}} D_{i+\frac{1}{2}}}{D_{i-\frac{1}{2}} + D_{i+\frac{1}{2}}}
  \end{equation}
  (harmonic mean), and similarly for $D_{i+1}$.

  Rearranging terms to bring all unknowns $e_{j+\frac{1}{2}}^{n+1}$ to the left yields the tridiagonal system:
  \begin{equation}
    \boxed{
      a_{i+\frac{1}{2}}\, e_{i-\frac{1}{2}}^{n+1}
      + b_{i+\frac{1}{2}}\, e_{i+\frac{1}{2}}^{n+1}
      + c_{i+\frac{1}{2}}\, e_{i+\frac{3}{2}}^{n+1}
      = e_{i+\frac{1}{2}}^n
    }
  \end{equation}
  where the boxed coefficients are:
  \begin{gather}
    \boxed{
      a_{i+\frac{1}{2}} = -\frac{\Delta t}{\Delta m_{i}} \frac{\rho_{i+\frac{1}{2}}^*}{\Delta m_{i+\frac{1}{2}}} D_{i}
    } \\[1em]
    \boxed{
      b_{i+\frac{1}{2}} = 1 + \frac{\Delta t}{\Delta m_{i+\frac{1}{2}}} \frac{\rho_{i+\frac{1}{2}}^*}{\Delta m_{i+\frac{1}{2}}} ( D_{i} + D_{i+1} )
    } \\[1em]
    \boxed{
      c_{i+\frac{1}{2}} = -\frac{\Delta t}{\Delta m_{i+1}} \frac{\rho_{i+\frac{1}{2}}^*}{\Delta m_{i+\frac{1}{2}}} D_{i+1}
    }
  \end{gather}

  Here, the index $i+\frac{1}{2}$ denotes the cell center, the left neighbor is at $i-\frac{1}{2}$ and the right neighbor at $i+\frac{3}{2}$, consistent with the Lagrangian grid convention.

  For practical implementation, let $j = i+\frac{1}{2}$ and define arrays for $a_j$, $b_j$, $c_j$, $e_j$, etc. The resulting tridiagonal system is:
  \begin{equation}
    a_j\, e_{j-1}^{n+1} + b_j\, e_j^{n+1} + c_j\, e_{j+1}^{n+1} = e_j^n
  \end{equation}
  which can be efficiently solved using the Thomas algorithm or any suitable tridiagonal matrix solver.


  Here, $e_{i+\frac{1}{2}}^{n+1}$ is the energy at cell center $i+\frac{1}{2}$ at the new timestep, and the resulting system is tridiagonal and can be solved by the Thomas algorithm.

  \textcolor{blue}{Note: If $e$ and $E$ are directly connected via LTE (e.g.\ $e=fT^\beta \rho^{-\mu}$ and $E=aT^4$) and the same equation holds for $E$, a similar discrete scheme applies to $E$.}
\end{enumerate}

\section{Appendix: Derivation of the tridiagonal form coefficients in Lagrangian coordinates}
\begin{enumerate}
  \item The discretized equation for the radiation energy in Lagrangian coordinates is:
  \begin{equation}
  \frac{E_{i+\frac{1}{2}}^{n+1} - E_{i+\frac{1}{2}}^n}{\Delta t} + \rho_{i+\frac{1}{2}}^{*} \frac{F_{i+1}^{n+1} - F_{i}^{n+1}}{\Delta m_{i+\frac{1}{2}}} =  \chi c \sigma_{i+\frac{1}{2}}^{n} ([U_R]_{i+\frac{1}{2}}^{n+1} - E_{i+\frac{1}{2}}^{n+1})
  \end{equation}
  \item Recalling that the flux and $U_R$ are given as expressions of $E$, we can substitute them into the equation above to get a tridiagonal system for $E_{i+\frac{1}{2}}^{n+1}$. \\ \\ 
  \underline{LHS:}
  The flux is given by:
  \begin{equation}
    F_{i}^{n} = -\rho_{i}^* D_{i}^n \frac{E_{i+\frac{1}{2}}^{n} - E_{i-\frac{1}{2}}^{n}}{\Delta m_{i}} 
  \end{equation}
  NOTE: The radiation energy density $E$ is taken implicitely, and the D face is taken exaplicitly - because of numeric considerations (\textcolor{blue}{I need to ask Shay about this}).
  Substituting the fluxes into the balance equation gives:
  \begin{equation}
    \begin{gathered}
  LHS = \frac{E_{i+\frac{1}{2}}^{n+1} - E_{i+\frac{1}{2}}^n}{\Delta t} - \frac{\rho_{i+\frac{1}{2}}^{*}}{\Delta m_{i+\frac{1}{2}}} \overbrace{\left(\rho_{i+1}^* D_{i+1}^n\frac{E_{i+\frac{3}{2}}^{n+1} - E_{i+\frac{1}{2}}^{n+1}}{\Delta m_{i+1}}\right)}^{-F_{i+1}^{n+1}} \\ + \frac{\rho_{i+\frac{1}{2}}^{*}}{\Delta m_{i+\frac{1}{2}}}\overbrace{\left(\rho_{i}^* D_{i}^n\frac{E_{i+\frac{1}{2}}^{n+1} - E_{i-\frac{1}{2}}^{n+1}}{\Delta m_{i}}\right)}^{-F_{i}^{n+1}}
    \end{gathered}
  \end{equation}
  Rearranging the terms gives us the tridiagonal form:
  \begin{equation}
    \begin{gathered}
      % insert the appropriate coefficients in front of each term:
  LHS = \left(-\frac{\rho_{i+\frac{1}{2}}^{*} \rho_i^* D_i^n}{\Delta m_{i+\frac{1}{2}} \Delta m_{i}} \right) E_{i-\frac{1}{2}}^{n+1} \\
  + \left[\frac{\rho_{i+\frac{1}{2}}^{*}}{\Delta m_{i+\frac{1}{2}}}\left(\frac{ \rho_{i}^* D_{i}^n}{ \Delta m_{i}}+\frac{ \rho_{i+1}^* D_{i+1}^n}{ \Delta m_{i+1}}\right) + \frac{1}{\Delta t}\right] E_{i+\frac{1}{2}}^{n+1} \\
  - \left(\frac{\rho_{i+\frac{1}{2}}^{*} \rho_{i+1}^* D_{i+1}^n}{\Delta m_{i+\frac{1}{2}} \Delta m_{i+1}}\right) E_{i+\frac{3}{2}}^{n+1} \\
  - \frac{1}{\Delta t} E_{i+\frac{1}{2}}^n
  \end{gathered}
  \end{equation}
  \underline{RHS:}
  Substituting the expression for $[U_R]_{i+\frac{1}{2}}^{n+1}$ gives:
  \begin{equation}
    \begin{gathered}
  RHS = \chi c \sigma_{i+\frac{1}{2}}^{n} \left(\frac{A_{i+\frac{1}{2}}^n}{1 + A_{i+\frac{1}{2}}^n} E_{i+\frac{1}{2}}^{n+1} + \frac{1}{1 + A_{i+\frac{1}{2}}^n}[U_R]_{i+\frac{1}{2}}^n - E_{i+\frac{1}{2}}^{n+1}\right) \\
  = \chi c \sigma_{i+\frac{1}{2}}^{n}
  \left(\frac{1}{1 + A_{i+\frac{1}{2}}^n}[U_R]_{i+\frac{1}{2}}^n -\frac{1}{1 + A_{i+\frac{1}{2}}^n} E_{i+\frac{1}{2}}^{n+1}\right) \\
  = \frac{\chi c \sigma_{i+\frac{1}{2}}^{n}}{1 + A_{i+\frac{1}{2}}^n}[U_R]_{i+\frac{1}{2}}^n - \frac{\chi c \sigma_{i+\frac{1}{2}}^{n}}{1 + A_{i+\frac{1}{2}}^n} E_{i+\frac{1}{2}}^{n+1} 
  \end{gathered}
  \end{equation}
  \item Finally, rearranging the equation $LHS = RHS$ gives us the tridiagonal form:
  \begin{equation}
    \begin{gathered}
     a_{i+\frac{1}{2}}^n E_{i-\frac{1}{2}}^{n+1} + b_{i+\frac{1}{2}}^n E_{i+\frac{1}{2}}^{n+1} + c_{i+\frac{1}{2}}^n E_{i+\frac{3}{2}}^{n+1} = d_{i+\frac{1}{2}}^n
    \end{gathered}
  \end{equation}
  and the coefficients are given by:
  \begin{gather}
    % add boxes around the coefficients:
    \boxed{a_{i+\frac{1}{2}}^n = -\frac{\rho_{i+\frac{1}{2}}^{*} \rho_i^* D_i^n}{\Delta m_{i+\frac{1}{2}} \Delta m_{i}}} \\
    \boxed{b_{i+\frac{1}{2}}^n = \frac{\rho_{i+\frac{1}{2}}^{*}}{\Delta m_{i+\frac{1}{2}}}\left(\frac{ \rho_{i}^* D_{i}^n}{ \Delta m_{i}}+\frac{ \rho_{i+1}^* D_{i+1}^n}{ \Delta m_{i+1}}\right) + \frac{1}{\Delta t} + \frac{\chi c \sigma_{i+\frac{1}{2}}^{n}}{1 + A_{i+\frac{1}{2}}^n}} \\
    \boxed{c_{i+\frac{1}{2}}^n = -\frac{\rho_{i+\frac{1}{2}}^{*} \rho_{i+1}^* D_{i+1}^n}{\Delta m_{i+\frac{1}{2}} \Delta m_{i+1}}} \\
    \boxed{d_{i+\frac{1}{2}}^n = \frac{\chi c \sigma_{i+\frac{1}{2}}^{n}}{1 + A_{i+\frac{1}{2}}^n}[U_R]_{i+\frac{1}{2}}^n + \frac{1}{\Delta t}E_{i+\frac{1}{2}}^n}
  \end{gather}
  where $A_{i+\frac{1}{2}}^n = \beta_{i+\frac{1}{2}}^{n} \Delta t \chi c \sigma_{i+\frac{1}{2}}^{n}$ is the coupling coefficient between the radiation and matter.
  \begin{notebox}[Two things to understand][red]
    I need to understand how the Flek \& Cummings coefficient $f$ is related to the coupling coefficient $A$.
  \end{notebox}
\end{enumerate}
  \subsection{Deeper understanding of the tridiagonal form}
  \begin{enumerate}
    \item First, let's make order in the celebration of indexes. Our staggered grid is as follows:
    \begin{center}
      \small
      \begin{tabular}{|c|c|c|c|c|c|c|}
      \hline
      \textbf{Start BC} & & \textbf{Internal} & $\dots$ & \textbf{Internal} & & \textbf{End BC} \\ \hline
      $i=0$ & $i=\frac{1}{2}$ & $i=1$ & $\dots$ & $i=N-1$ & $i=N-\frac{1}{2}$ & $i=N$ \\ 
      $j=\frac{1}{2}$ & $\boxed{j=1}$ & $j=\frac{3}{2}$ & $\dots$ & $j=N-\frac{1}{2}$ & $\boxed{j=N}$ & $j=N+\frac{1}{2}$ \\ \hline
      node & cell center & node & $\dots$ & node & cell center & node \\ 
      $\rho_0, D_0$ & $E_0, [U_R]_0$ & $\rho_1, D_1$ & $\dots$ & $\rho_{N-1}, D_{N-1}$ & $E_N, [U_R]_N$ & $\rho_N, D_N$ \\ 
      & $\rho_{1/2}, D_{1/2}$ & & $\dots$ & & $\rho_{N-1/2}, D_{N-1/2}$ & \\ 
      & $\beta_{1/2}, \sigma_{1/2}$ & & $\dots$ & & $\beta_{N-1/2}, \sigma_{N-1/2}$ & \\ \hline
      \end{tabular}
  \end{center}
    where we defined $j=i + \frac{1}{2}$ as the index for the cell centers. Now, the indexes values are:
    \begin{equation}
      \begin{gathered}
        i = 0,1,2,...,N \quad (\text{nodes}) \\
        j = 1,2,...,N \quad (\text{cell centers})
      \end{gathered}
    \end{equation}
    \item We wish to write the tridiagonal equations in a comfortable matrix form. Under the assumption that the radiation energy array $E_j$ is defined on cell centers $j = 1,...,N$ but also at $j=0$ and $j=N+1$ (ghost cells in which the value is defined via the BC), we get the follows. \newline
    \underline{For $j = 1, ..., N$ where $i = j - \frac{1}{2}$}: 
    \begin{equation}
      \begin{gathered}
        a_{i+\frac{1}{2}}^n E_{i-\frac{1}{2}}^{n+1} + b_{i+\frac{1}{2}}^n E_{i+\frac{1}{2}}^{n+1} + c_{i+\frac{1}{2}}^n E_{i+\frac{3}{2}}^{n+1} = d_{i+\frac{1}{2}}^n \\
        \Downarrow\\
        a_{j}^n E_{j-1}^{n+1} + b_{j}^n E_{j}^{n+1} + c_{j}^n E_{j+1}^{n+1} = d_{j}^n \\
      \end{gathered}
    \end{equation}
    which build the following matrix system of equations:
    \begin{equation}
    \begin{bmatrix}
    a_1 & b_1 & c_1 & 0 & \cdots & 0 & 0\\
    0 & a_2 & b_2 & c_2 & \cdots & 0 & 0 \\  
    0 & 0 & a_3 & b_3 & c_3 &\cdots & 0 \\
    \vdots & \vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
    0  & \cdots & 0 & 0 & a_{N-1} & b_{N-1} & c_{N-1} & 0 \\
    0 & 0 & 0 & \cdots & 0 & a_{N} & b_{N} & c_{N}
    \end{bmatrix}
    \begin{bmatrix}
    E_0^{n+1} \\
    E_1^{n+1} \\
    \vdots \\
    \vdots \\
    \vdots \\
    E_{N}^{n+1} \\
    E_{N+1}^{n+1}
    \end{bmatrix}
    =
    \begin{bmatrix}
    d_1 \\
    d_2 \\
    d_3 \\
    \vdots \\
    d_{N-1} \\
    d_{N}
    \end{bmatrix}
    \end{equation}
    which gives us a system of \textbf{$N$ equations with $N+2$ variables}. \newline \newline
    Before we discuss the definition of the cell-centered coefficients $a_j^n, b_j^n, c_j^n, d_j^n$ for $j=1,...,N$, let's think of how one would implement them on a \textbf{computer program}.
    \begin{notebox}[Implementation of the coefficient calculations][blue]
      There are two important factors to consider when calculating the coefficients:
      \begin{enumerate}
        \item How one correctly implement the cell-valued arrays and the face-valued arrays in a computer program, where there isn't a way to create half indexes or differentiate between $i$ and $j$ in a simple array.
        \item The validity of the indexes of the coefficients and other variables are affected by the boundary conditions.
      \end{enumerate}
    \end{notebox}
    \subsubsection{The matrix equation for different BCs}
    \begin{enumerate}
    \item In \textbf{Dirichlet BC} we define:
    \begin{gather}
      E_0^{n} = T_s(t) \\
      E_{N+1}^{n} = 0
    \end{gather}
    so that the system of equations is now:
    \begin{equation}
    \begin{bmatrix}
    a_1 & b_1 & c_1 & 0 & \cdots & 0 & 0\\
    0 & a_2 & b_2 & c_2 & \cdots & 0 & 0 \\  
    0 & 0 & a_3 & b_3 & c_3 &\cdots & 0 \\
    \vdots & \vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
    0  & \cdots & 0 & 0 & a_{N-1} & b_{N-1} & c_{N-1} & 0 \\
    0 & 0 & 0 & \cdots & 0 & a_{N} & b_{N} & c_{N}
    \end{bmatrix}
    \begin{bmatrix}
    T_s(t) \\
    E_1^{n+1} \\
    \vdots \\
    \vdots \\
    \vdots \\
    E_{N}^{n+1} \\
    0
    \end{bmatrix}
    =
    \begin{bmatrix}
    d_1 \\
    d_2 \\
    d_3 \\
    \vdots \\
    d_{N-1} \\
    d_{N}
    \end{bmatrix}
    \end{equation}
    And now we are left with $N$ equations with $N$ unknowns! (in fact we have just added the two missing equations, that directly solve for $E_0$ and $E_{N+1}$). \newline \newline
    At this point, the $1$st and $N$'th equations become:
    \begin{gather}
      a_1 T_s(t) + b_1 E_1^{n+1} + c_1 E_2^{n+1} = d_1 \\
      a_N E_{N-1}^{n+1} + b_N E_N^{n+1} + c_N \cdot 0= d_N
    \end{gather}
    which can be written as:
    \begin{gather}
      b_1 E_1^{n+1} + c_1 E_2^{n+1} = d_1 - a_1 T_s(t)\\
      a_N E_{N-1}^{n+1} + b_N E_N^{n+1} = d_N
    \end{gather}
    So that the elements $a_1$ and $c_N$ are ommitted, and their effect is inserted inside the $d$ coefficients to provide the \textbf{reduced matrix equation:}
    \begin{equation}
    \begin{bmatrix}
    b_1 & c_1 & 0 & \cdots & 0 \\
    a_2 & b_2 & c_2 & \cdots & 0  \\  
    0 & a_3 & b_3 & c_3 &\cdots  \\
    \vdots & \vdots & \vdots & \ddots & \vdots  \\
    \cdots & 0 & 0 & a_{N-1} & b_{N-1} & c_{N-1} \\
    0 & 0 & \cdots & 0 & a_{N} & b_{N}
    \end{bmatrix}
    \begin{bmatrix}
    E_1^{n+1} \\
    \vdots \\
    \vdots \\
    \vdots \\
    E_{N}^{n+1} \\
    \end{bmatrix}
    =
    \begin{bmatrix}
    d_1 - a_1 T_s(t) \\
    d_2 \\
    d_3 \\
    \vdots \\
    d_{N-1} \\
    d_{N}
    \end{bmatrix}
    \end{equation}
    \item In \textbf{Marshak BC} we define:
    \begin{gather}
      F_0^{n} = F(T_{bath}) \\
      E_{N+1}^{n} = 0
    \end{gather}
    Through a derivation that appears in detail in Appendix 6 % make that clickable
    one can reduce the equation at the surface ($i=0$) into an equation of the form:
    \begin{equation}
      \tilde{b}_1 E_1^{n+1} + \tilde{c}_1 E_2^{n+1} = \tilde{d}_1 
    \end{equation}
    where $\tilde{b}_1, \tilde{c}_1, \tilde{d}_1$ are of course defined in Appendix 6.
    Therefore, after ommiting the last equation (which involves a trivial Dirichlet BC at the edge ghost cell):
    \begin{equation}
    \begin{bmatrix}
    0 & \tilde{b}_1 & \tilde{c}_1 & 0 & \cdots & 0 & 0\\
    a_1 & b_1 & c_1 & 0 & \cdots & 0 & 0\\
    0 & a_2 & b_2 & c_2 & \cdots & 0 & 0 \\  
    0 & 0 & a_3 & b_3 & c_3 &\cdots & 0 \\
    \vdots & \vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
    0  & \cdots & 0 & 0 & a_{N-1} & b_{N-1} & c_{N-1} & 0 \\
    0 & 0 & 0 & \cdots & 0 & a_{N} & b_{N} & c_{N}
    \end{bmatrix}
    \begin{bmatrix}
    E_0^{n+1} \\
    E_1^{n+1} \\
    \vdots \\
    \vdots \\
    \vdots \\
    E_{N}^{n+1} \\
    \end{bmatrix}
    =
    \begin{bmatrix}
    \tilde{d}_1 \\
    d_1 \\
    d_2 \\
    d_3 \\
    \vdots \\
    d_{N-1} \\
    d_{N}
    \end{bmatrix}
    \end{equation}
    We are now left with $N+1$ equations in $N+1$ unknowns.
    \end{enumerate}
    \subsubsection{The tridiagonal coefficient calculation}
    $\rho, \Delta m, D, \sigma$ are all given as arrays for the cell centers (i.e. $\rho_j$ = \texttt{rho[j - 1]}) for $j = 1,...,N$. Starting from this principle, we can calculate the face-valued arrays as follows. 
    \newline \underline{For $i = 0,..., N$} (notice the \texttt{[1:], [:-1]} notation that forbids invalid indices):
    \begin{equation}
      \begin{gathered}
        \rho_i^* = \text{avg}(\rho_{i-\frac{1}{2}}^*, \rho_{i+\frac{1}{2}}^*) \implies \texttt{rho\_face[:]=avg(rho[:-1], rho[1:])}\\
        D_i^n = \text{avg}(D_{i-\frac{1}{2}}^n, D_{i+\frac{1}{2}}^n) \implies \texttt{D\_face[:]=avg(D[:-1], D[1:])}\\
        \Delta m_i = \text{avg}(\Delta m_{i-\frac{1}{2}}, \Delta m_{i+\frac{1}{2}}) \implies \texttt{m\_face[:]=avg(m\_cells[:-1], m\_cells[1:])}\\
      \end{gathered}
    \end{equation}
    of \texttt{size=N-1}.
    We will define:
    \begin{equation}
      f_i^* = \frac{\rho_i^* D_i^n}{\Delta m_i} 
    \end{equation}
    and now we are ready to define the \texttt{flux\_coeff} so that \texttt{flux\_coeff.size()} is $N-1$:
    \begin{equation}
      \texttt{flux\_coeff}[i - 1] = f_{i}^*
    \end{equation}
    \begin{notebox}[Face values at the first node ($i=0$)]
      Since the size of the \texttt{flux\_coeff} array is $N-1$, and it should be multiplied element-wise by other $N$ sized arrays (e.g., \texttt{lagrangian\_coeff}) that are computed at the cell centers (specifically - the cell center of which the corresponding node is located at its left face since $i = j - 1/2$), we shall define the value of $\rho^*$, $D^n$, $\Delta m$ at $i=0$ by reflecting their values across the boundary, i.e., $f_0^* = f_1^*$.
    \end{notebox}
    \item We will also define the following numeric arrays for notation comfort.
    \newline \underline{For $j = 1,...,N$}: 
    Under the definitions:
    \begin{equation}
      \begin{gathered}
        \mathcal{L}_j^* = \frac{\rho_j^*}{\Delta m_j} \\
        B_j^n = \frac{\chi c \sigma_j^n}{1 + A_j^n}
      \end{gathered}
    \end{equation}
    The numeric arrays are:
    \begin{equation}
      \begin{gathered}
        \texttt{lagrangian\_coeff[j-1]} = \mathcal{L}_j^*\\
        \texttt{total\_coupling[j-1]} = B_j^n
      \end{gathered}
    \end{equation}
    \item To sum up, the coefficients are defined as follows. \newline
    \underline{For $j = 1,...,N$ and $i = j $ (left face)}:
    \begin{equation}
      \begin{gathered}
        a_{j}^n = -\mathcal{L}_j^* f_i^* \\
        b_{j}^n = \mathcal{L}_j^*(f_i^* + f_{i+1}^*) + \frac{1}{\Delta t} + B_j^n \\
        c_{j}^n = -\mathcal{L}_j^* f_{i+1}^* \\
        d_{j}^n = B_j^n[U_R]_{j}^n + \frac{1}{\Delta t}E_{j}^n
      \end{gathered}
    \end{equation}
    and in code notation (after defining \texttt{a, b, c, d} arrays of \texttt{size = N}):
    \begin{equation}
      \begin{gathered}
        \texttt{a[j - 1]} = \mathcal{L}_j^* f_i^* \\
        \texttt{b[j - 1]} = \mathcal{L}_j^*(f_i^* + f_{i+1}^*) + \frac{1}{\Delta t} + B_j^n \\
        \texttt{c[j - 1]} = \mathcal{L}_j^* f_{i+1}^* \\
        \texttt{d[j - 1]} = B_j^n[U_R]_{j}^n + \frac{1}{\Delta t}E_{j}^n
      \end{gathered}
    \end{equation}
    \begin{notebox}[Between code arrays and physical quantities]
      As stated at the beginning of this subsection, the physical quantities $\rho, \Delta m, D, \sigma$ are all given as arrays for the cell centers (i.e. $\rho_j$ = \texttt{rho[j - 1]}) for $j = 1,...,N$ so assumptions such as \texttt{arr[j-1]} = $f(\rho, \Delta m, D, \sigma)$ just mean direct array assignment!
    \end{notebox}
  \end{enumerate}


  \newpage
  \newpage
  \section{Appendix: Boundary Conditions in the Diffusion Approximation}
  
  In the $P_1$ diffusion approximation, the specific intensity $I(x, \mu)$ is not tracked for every angle. Instead, we use \textbf{Marshak Boundary Conditions} to bridge the gap between averaged quantities (energy density $E$ and flux $F$) and the physical constraints at the boundaries.
  
\subsection{General Physical Derivation}
At a boundary interface, the net inward partial flux $J_{in}$ must match the source drive. For a blackbody source at temperature $T_{bath}$, $J_{in} = \sigma_{sb} T_{bath}^4$. In the diffusion approximation, the partial fluxes are related to $E$ and $F$ by:
\begin{equation}
    J_{in} = \frac{cE}{4} + \frac{F}{2}, \quad J_{out} = \frac{cE}{4} - \frac{F}{2}
\end{equation}
At the origin ($x=0$), the net flux into the material is $F_0$. By substituting $\sigma_{sb} T_{bath}^4 = \frac{c}{4} E_{bath}$, the Marshak condition becomes:
\begin{gather}
    \frac{cE(0,t)}{4} + \frac{F_0(t)}{2} = \frac{c}{4} E_{bath}(t) \\
    \boxed{F_0(t) = \frac{c}{2} E_{bath}(t) - \frac{c}{2} E(0,t)}
\end{gather}
At the far boundary ($x=L$), we assume a vacuum where no radiation enters from the right ($J_{in, right} = 0$), leading to the free-streaming limit:
\begin{equation}
    \frac{cE(L,t)}{4} - \frac{F_L(t)}{2} = 0 \implies \boxed{F_L = \frac{cE(L,t)}{2}}
\end{equation}

\subsection{Discretization in Lagrangian Coordinates}
In our staggered grid, cell centers are indexed by $j=1, \dots, N$ and faces by $i=0, \dots, N$. The diffusion equation at the cell located between the faces $i$ and $i+1$ is:
\begin{equation}
\begin{gathered}
    \frac{E_{i+\frac{1}{2}}^{n+1} - E_{i+\frac{1}{2}}^n}{\Delta t} - \frac{\rho_{i+\frac{1}{2}}^{*}}{\Delta m_{i+\frac{1}{2}}} \overbrace{\left(\rho_{i+1}^n D_{i+1}^n\frac{E_{i+\frac{3}{2}}^{n+1} - E_{i+\frac{1}{2}}^{n+1}}{\Delta m_{i+1}}\right)}^{-\hat{F}_{i+1}^{n+1}} + \frac{\rho_{i+\frac{1}{2}}^{*}}{\Delta m_{i+\frac{1}{2}}}\overbrace{\left(\rho_{i}^n D_{i}^n\frac{E_{i+\frac{1}{2}}^{n+1} - E_{i-\frac{1}{2}}^{n+1}}{\Delta m_{i}}\right)}^{-\hat{F}_{i}^{n+1}} = \\ 
    \frac{\chi c \sigma_{i+\frac{1}{2}}^{n}}{1 + A_{i+\frac{1}{2}}^n}[U_R]_{i+\frac{1}{2}}^n - \frac{\chi c \sigma_{i+\frac{1}{2}}^{n}}{1 + A_{i+\frac{1}{2}}^n} E_{i+\frac{1}{2}}^{n+1} 
\end{gathered}   
\end{equation}
\begin{notebox}[Quick note about notation]
I used the notation "$\hat{F}_1$" to indicate that the subscript denotes $i$ and not $j$. For comfort indexing, in the notation "$E_1$" (without hat) the subscript denotes $j$ (default notation prefers cell centers).
\end{notebox}
\subsubsection*{First Cell ($j=1$)}
For the first cell, we substitute the boundary flux $\hat{F}_0^{n+1}$ derived from the Marshak condition:
\begin{equation}
    \frac{E_{1}^{n+1} - E_{1}^n}{\Delta t} + \frac{\rho_{1}^{*}}{\Delta m_{1}} \left( \hat{F}_{1}^{n+1} - \overbrace{\left[ \frac{c}{2} E_{bath} - \frac{c}{2} E(0,t) \right]}^{\hat{F}_0^{n+1}} \right) = S_1^{n+1}
\end{equation}
where the source term is defined as:
\begin{equation}
    S_1^{n+1} = \frac{\chi c \sigma_1^n}{1 + A_1^n}[U_R]_1^n - \frac{\chi c \sigma_1^n}{1 + A_1^n} E_1^{n+1}
\end{equation}

To close the system using only internal cell unknowns, we make the standard first-order spatial approximation that the radiation energy density at the boundary face is roughly equal to the energy of the first cell: $E(0,t) \approx E_1^{n+1}$. Substituting this and the internal flux $\hat{F}_1^{n+1}$ yields:
\begin{equation}
\begin{aligned}
    \frac{E_1^{n+1} - E_1^n}{\Delta t} + \frac{\rho_1^*}{\Delta m_1} & \left[ \left( -\hat{\rho}_1^n \hat{D}_1^n \frac{E_2^{n+1} - E_1^{n+1}}{\Delta \hat{m}_1} \right) \right. \\
    & \left. - \frac{c}{2} E_{bath} + \frac{c}{2} E_1^{n+1} \right] = \\
    & \frac{\chi c \sigma_1^n}{1 + A_1^n} [U_R]_1^n - \frac{\chi c \sigma_1^n}{1 + A_1^n} E_1^{n+1}
\end{aligned}
\end{equation}

Notice that the surface emission term has become $\frac{\rho_1^* c}{2\Delta m_1} E_1^{n+1}$. This naturally moves to the LHS of the equation to maintain the implicit scheme stability. Let us define this Marshak boundary cooling coefficient $C_1$:
\begin{equation}
    C_1 = \frac{c \rho_1^*}{2 \Delta m_1}
\end{equation}

Grouping terms for $E_1^{n+1}$ and $E_2^{n+1}$ yields the modified tridiagonal system:
\begin{equation}
\begin{aligned}
    & \left( \frac{1}{\Delta t} + \frac{\rho_1^* \hat{\rho}_1^n \hat{D}_1^n}{\Delta m_1 \Delta \hat{m}_1} + C_1 + \frac{\chi c \sigma_1^n}{1 + A_1^n} \right)E_1^{n+1} \\
    & - \left( \frac{\rho_1^* \hat{\rho}_1^n \hat{D}_1^n}{\Delta m_1 \Delta \hat{m}_1} \right)E_2^{n+1} = \\
    & \frac{E_1^n}{\Delta t} + C_1 E_{bath} + \frac{\chi c \sigma_1^n}{1 + A_1^n} [U_R]_1^n
\end{aligned}
\end{equation}
where $E_{bath} = a T_{bath}^4(t)$. 

This gives us the modified tridiagonal coefficients for the first cell:
\begin{gather}
    \boxed{ \tilde{b}_1^n = \frac{1}{\Delta t} + \frac{\rho_1^* \hat{\rho}_1^n \hat{D}_1^n}{\Delta m_1 \Delta \hat{m}_1} + C_1 + \frac{\chi c \sigma_{1}^{n}}{1 + A_{1}^n} } \\
    \boxed{ \tilde{c}_1^n = -\frac{\rho_1^* \hat{\rho}_1^n \hat{D}_1^n}{\Delta m_1 \Delta \hat{m}_1} } \\
    \boxed{ \tilde{d}_1^n = \frac{E_1^n}{\Delta t} + C_1 E_{bath} + \frac{\chi c \sigma_{1}^{n}}{1 + A_{1}^n} [U_R]_{1}^n }
\end{gather}

\begin{notebox}[Relation to the base internal coefficients][blue]
We can connect these modified coefficients to the base internal coefficients $a_1^n, b_1^n, d_1^n$ derived earlier. Recall that $a_1^n$ represents the left-face diffusion contribution, which is $-\frac{\rho_1^* \hat{\rho}_0^n \hat{D}_0^n}{\Delta m_1 \Delta \hat{m}_0}$, and $b_1^n$ contains $-a_1^n$. The Marshak condition essentially replaces this left-face diffusion with the Marshak cooling/drive terms. 

Therefore, the modified coefficients can be elegantly constructed directly from the base coefficients:
\begin{gather}
    \boxed{ \tilde{b}_1^n = b_1^n + a_1^n + C_1 } \\
    \boxed{ \tilde{a}_1^n = 0 } \\
    \boxed{ \tilde{d}_1^n = d_1^n + C_1 E_{bath} }
\end{gather}
This avoids recalculating the shared terms (like the right-face flux or the source terms) and is exactly how it is implemented numerically.
\end{notebox}
  \paragraph{Last Cell ($i=N-\frac{1}{2}, j=N$):}
  This is not done in our simulation because the Marshak waves don't reach the edge boundary in our simulation setup.
  \begin{equation}
      \frac{E_N^{n+1} - E_N^n}{\Delta t} + \frac{\rho_N^*}{\Delta m_N} \left( \frac{c E_N^{n+1}}{2} - F_{N-1}^{n+1} \right) = S_N^{n+1}
  \end{equation}
  This modifies the final diagonal element:
  \begin{equation}
      \boxed{\tilde{b}_N^n = b_N^n + \frac{c \rho_N^*}{2 \Delta m_N}}
  \end{equation}
\end{document}